---
title: "Machine Learning for Business Analytics"
author: "Chapter 10: Logistic Regression"
output:
  pdf_document:
    toc: no
    highlight: tango
#  html_document:
#    toc: yes
#    toc_depth: 4
#    toc_float: yes
---
<style>
h1.title { font-size: 28px; }
h1 { font-size: 22px; }
h2 { font-size: 18px; }
h3 { font-size: 14px; }
h4 { font-size: 12px; }
</style>
```{r global_options, include=FALSE}
knitr::opts_chunk$set(error=TRUE,        # Keep compiling upon error
                      collapse=FALSE,    # collapse by default
                      echo=TRUE,         # echo code by default
                      comment = "#>",    # change comment character
                      fig.width = 5.5,     # set figure width
                      fig.align = "center",# set figure position
                      out.width = "49%", # set width of displayed images
                      warning=FALSE,     # do not show R warnings
                      message=FALSE,     # do not show R messages
                      row.print=25)
```


```{r}
options(scipen=999)

# TODO: remove once package is available on CRAN
library(devtools)
install_github("gedeck/mlba/mlba", force=TRUE)
# library(mlba)
```

# The Logistic Regression Model
```{r}
library(ggplot2)
  library(gridExtra)

  p <- seq(0.005, 0.995, 0.01)
  df <- data.frame(
    p = p,
    odds = p / (1 - p),
    logit = log(p / (1 - p))
  )

  g1 <- ggplot(df, aes(x=p, y=odds)) +
    geom_line() + coord_cartesian(xlim=c(0, 1), ylim=c(0, 100)) +
    labs(x='Probability of success', y='Odds', title='(a)') +
    geom_hline(yintercept = 0) +
    theme_bw() +
    theme(axis.line = element_line(colour = "black"),
          axis.line.x = element_blank(),
          panel.border = element_blank())
  g2 <- ggplot(df, aes(x=p, y=logit)) +
    geom_line() + coord_cartesian(xlim=c(0, 1), ylim=c(-4, 4)) +
    labs(x='Probability of success', y='Logit', title='(b)' ) +
    geom_hline(yintercept = 0) +
    theme_bw() +
    theme(axis.line = element_line(colour = "black"),
          axis.line.x = element_blank(),
          panel.border = element_blank())

  grid.arrange(g1, g2, ncol=2)

  g1 <- ggplotGrob(g1)
  g2 <- ggplotGrob(g2)
  maxWidth = grid::unit.pmax(g1$widths[2:5], g2$widths[2:5])
  g1$widths[2:5] <- as.list(maxWidth)
  g2$widths[2:5] <- as.list(maxWidth)
  g <- arrangeGrob(g1, g2)
  ggsave(file=file.path("..", "figures", "chapter_10", "odds_logit.pdf"), g,
         width=6, height=6)
```

# Example: Acceptance of Personal Loan
## Model with a Single Predictor
```{r}
bank.df <- mlba::UniversalBank
g <- ggplot(bank.df, aes(x=Income, y=Personal.Loan)) +
  geom_jitter(width=0, height=0.01, alpha=0.1) +
  geom_function(fun=function(x){ return (1 / (1 + exp(6.04892 - 0.036*x)))}) +
  xlim(0, 250) +
  labs(x='Income (in $000s)') +
  theme_bw()
g
ggsave(file=file.path("..", "figures", "chapter_10", "fitted_logistic.pdf"), g,
         width=5, height=3)
```

## Estimating the Logistic Model from Data: Computing Parameter Estimates
### Estimated Model
```{r}
library(caret)
library(tidyverse)

# load and preprocess data
bank.df <- mlba::UniversalBank %>%
  select(-c(ID, ZIP.Code)) %>% # Drop ID and zip code columns.
  mutate(
    Education = factor(Education, levels=c(1:3),
                       labels=c("Undergrad", "Graduate", "Advanced/Professional")),
    Personal.Loan = factor(Personal.Loan, levels=c(0, 1),
                           labels=c("No", "Yes"))
  )

# partition data
set.seed(2)
idx <- caret::createDataPartition(bank.df$Personal.Loan, p=0.6, list=FALSE)
train.df <- bank.df[idx, ]
holdout.df <- bank.df[-idx, ]

# build model
trControl <- caret::trainControl(method="cv", number=5, allowParallel=TRUE)
logit.reg <- caret::train(Personal.Loan ~ ., data=train.df, trControl=trControl,
                      # fit logistic regression with a generalized linear model
                      method="glm", family="binomial")
logit.reg
summary(logit.reg$finalModel)
```

# Evaluating Classification Performance
## Interpreting Results in Terms of Odds (for a Profiling Goal)
```{r}
# use predict() with type = "response" to compute predicted probabilities.
logit.reg.pred <- predict(logit.reg, holdout.df[, -8], type = "prob")

# display four different cases
interestingCases = c(1, 12, 32, 1333)
data.frame(
  actual = holdout.df$Personal.Loan[interestingCases],
  p0 = logit.reg.pred[interestingCases, 1],
  p1 = logit.reg.pred[interestingCases, 2],
  predicted = ifelse(logit.reg.pred[interestingCases, 2] > 0.5, 1, 0)
)
```

```{r}
library(gains)
actual <- ifelse(holdout.df$Personal.Loan == "Yes", 1, 0)
gain <- gains(actual, logit.reg.pred[,2], groups=length(actual) - 2)

# plot gains chart
nactual <-sum(actual)
g1 <- ggplot() +
  geom_line(aes(x=gain$cume.obs, y=gain$cume.pct.of.total * nactual)) +
  geom_line(aes(x=c(0, max(gain$cume.obs)), y=c(0, nactual)), color="darkgrey") +
  labs(x="# cases", y="Cumulative")

# plot decile-wise lift chart
gain10 <- gains(actual, logit.reg.pred[,2], groups=10)
g2 <- ggplot(mapping=aes(x=gain10$depth, y=gain10$lift / 100)) +
  geom_col(fill="steelblue") +
  geom_text(aes(label=round(gain10$lift / 100, 1)), vjust=-0.2, size=3) +
  ylim(0, 8) + labs(x="Percentile", y="Lift")
grid.arrange(g1, g2, ncol=2)
```

```{r}
g <- arrangeGrob(g1 + theme_bw(), g2 + theme_bw(), ncol=2, widths=c(0.45, 0.55))
ggsave(file=file.path("..", "figures", "chapter_10", "gains-lift.pdf"), g,
       width=6, height=2.5)
```

# Logistic Regression for Multi-Class Classification
### Ordinal Classes
```{r}
# simulate simple data
Y = rep(c("a", "b", "c"), 100)
x = rep(c(1, 2, 3), 100) +  rnorm(300, 0, 1)

# ordinal logistic regression
Y = factor(Y, ordered = T)
MASS::polr(Y ~ x)

# nominal logistic regression
Y = factor(Y, ordered = F)
nnet::multinom(Y ~ x)
```

# Example of Complete Analysis: Predicting Delayed Flights
```{r}
# code for generating top-left bar chart
# for other plots replace the aggregating variable in the group_by and ggplot commands
# adjust the x-label
library(tidyverse)

delays.df <- mlba::FlightDelays
averageDelay <- delays.df %>%
  group_by(DAY_WEEK) %>%
  summarize(mean=mean(Flight.Status == 'delayed'))

ggplot(averageDelay, aes(x=DAY_WEEK, y=mean)) +
  geom_col(fill='steelblue') +
  labs(x='Day of Week', y='Average Delay')
```

```{r}
averageDelayChart <- function(data, groupBy, xlabel) {
  averageDelay <- data %>%
    mutate(groupBy = factor(groupBy)) %>%
    group_by_at(groupBy) %>%
    summarize(mean=mean(Flight.Status == 'delayed'))

  return (ggplot(averageDelay, aes_string(x=groupBy, y='mean')) +
    geom_col(fill='steelblue') +
    theme_bw() +
    labs(x=xlabel, y='Average Delay'))
}
binned.df <- data.frame(
  Flight.Status = delays.df$Flight.Status,
  CRS_DEP_TIME = round(delays.df$CRS_DEP_TIME / 100)
)
# need to change Weather to factor
delays.df$Weather <- factor(delays.df$Weather)
g1 <- averageDelayChart(delays.df, 'DAY_WEEK', 'Day of week')
g2 <- averageDelayChart(delays.df, 'DEST', 'Destination')
g3 <- averageDelayChart(binned.df, 'CRS_DEP_TIME', 'Departure time (planned)')
g4 <- averageDelayChart(delays.df, 'CARRIER', 'Carrier')
g5 <- averageDelayChart(delays.df, 'ORIGIN', 'Origin')
g6 <- averageDelayChart(delays.df, 'Weather', 'Weather')

grid.arrange(g1, g2, g3, g4, g5, g6, ncol=2, nrow=3)
g <- arrangeGrob(g1, g2, g3, g4, g5, g6, ncol=2, nrow=3)
ggsave(file=file.path("..", "figures", "chapter_10", "LRFlightDelaysBarCharts_new.pdf"),
       g, width=6, height=6, units="in")
```

```{r}
# code for generating top-right bar chart
# for other plots, replace aggregating variable by setting argument by =  in
# aggregate().
# in function barplot(), set the x-label (argument xlab =) and y-label
# (argument names.arg =)
# according to the variable of choice.
delays.df = mlba::FlightDelays
barplot(aggregate(delays.df$Flight.Status == "delayed", by = list(delays.df$DAY_WEEK),
                  mean, rm.na = T)[,2], xlab = "Day of Week", ylab = "Average Delay",
                  names.arg = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun"))
```

```{r}
library(reshape)
# create matrix for plot
agg <- delays.df %>%
  mutate(isDelay=1 * (delays.df$Flight.Status == "delayed")) %>%
  group_by(DAY_WEEK, CARRIER, ORIGIN) %>%
  summarize(mean=mean(isDelay))

# plot with ggplot
# use facet_grid() with arguments scales = "free" and space = "free" to skip
# missing values.
ggplot(agg, aes(y=CARRIER, x=DAY_WEEK, fill=mean)) +
  geom_tile() +
  facet_grid(ORIGIN ~ ., scales="free", space="free") +
  scale_fill_gradient(low="white", high="steelblue")
```

```{r}
ggsave(file=file.path("..", "figures", "chapter_10", "LRFlightDelaysHeatmap.pdf"),
         last_plot() + theme_bw(), width=6, height=4, units="in")
```

## Model-Fitting and Estimation
```{r}
df <- mlba::FlightDelays %>%
  mutate(
    # transform variables and create bins
    DAY_WEEK = factor(DAY_WEEK, levels=c(1:7),
                      labels=c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")),
    CRS_DEP_TIME = factor(round(CRS_DEP_TIME / 100)),
    Flight.Status = factor(Flight.Status, levels=c("ontime", "delayed")),
    # convert other variables to factors
    across(c(ORIGIN,DEST, CARRIER, DAY_WEEK), factor)
  ) %>%
  # select predictors and outcome variables
  select(DAY_WEEK, CRS_DEP_TIME, ORIGIN, DEST, CARRIER, Weather, Flight.Status)

# create training and holdout sets
set.seed(1)
idx <- caret::createDataPartition(df$Flight.Status, p=0.6, list=FALSE)
train.df <- df[idx, ]
holdout.df <- df[-idx, ]

# run logistic model, and show coefficients and odds
trControl <- caret::trainControl(method="cv", number=5, allowParallel=TRUE)
model <- caret::train(Flight.Status ~ ., data=train.df,
                      method="glm",  family="binomial", trControl=trControl)
round(data.frame(summary(model$finalModel)$coefficients,
                 odds = exp(coef(model$finalModel))), 5)
```

## Model Performance
```{r}
library(gains)
actual <- ifelse(holdout.df$Flight.Status == "delayed", 1, 0)
pred <- predict(model, holdout.df, type="prob")[, 2]
gain <- gains(actual, pred, groups=length(pred))

nactual <-sum(actual)
ggplot() +
  geom_line(aes(x=gain$cume.obs, y=gain$cume.pct.of.total * nactual)) +
  geom_line(aes(x=c(0, max(gain$cume.obs)), y=c(0, nactual)), color="darkgrey") +
  labs(x="# cases", y="Cumulative")
```

```{r}
confusionMatrix(predict(model, holdout.df), holdout.df$Flight.Status)
```

```{r}
ggsave(file=file.path("..", "figures", "chapter_10", "LRperformance_Flights.pdf"),
       last_plot() + theme_bw(),
       width=3, height=3)
```

## Variable Selection
```{r}
# fewer predictors
reduced.df <- mlba::FlightDelays %>%
  mutate(
    HOUR = round(CRS_DEP_TIME / 100),
    SUN_MON = DAY_WEEK %in% c(1, 7),
    CARRIER_CO_MQ = CARRIER %in% c("CO", "MQ"),
    CARRIER_DL_US = CARRIER %in% c("DL", "US"),
    CRS_DEP_TIME15 = HOUR %in% c(15),
    EVENING = HOUR %in% c(19, 20, 21),
    Flight.Status = factor(Flight.Status, levels=c("ontime", "delayed"))
  ) %>%
  select(Weather, SUN_MON, CARRIER_CO_MQ, CARRIER_DL_US, CRS_DEP_TIME15, EVENING, Flight.Status)
# create training and holdout sets
set.seed(1)
idx <- caret::createDataPartition(reduced.df$Flight.Status, p=0.6, list=FALSE)
train.df <- reduced.df[idx, ]
holdout.df <- reduced.df[-idx, ]

# run logistic model, and show coefficients and odds
model <- caret::train(Flight.Status ~ ., data=train.df,
                      method="glm", family="binomial", trControl=trControl)
summary(model$finalModel)

# evaluate
pred <- predict(model, holdout.df)
confusionMatrix(pred, holdout.df$Flight.Status)
```

```{r}
actual <- ifelse(holdout.df$Flight.Status == "delayed", 1, 0)
pred <- predict(model, holdout.df, type="prob")[, 2]
redGain <- gains(actual, pred, groups=length(pred))
```

```{r}
# recreate training and holdout sets
set.seed(1)
idx <- caret::createDataPartition(df$Flight.Status, p=0.6, list=FALSE)
train.df <- df[idx, ]
holdout.df <- df[-idx, ]

# run logistic model, and show coefficients and odds
trControl <- caret::trainControl(method="cv", number=5, allowParallel=TRUE)
tuneGrid <- expand.grid(lambda=10^seq(-1, -3, by=-0.1), alpha=1)
model <- caret::train(Flight.Status ~ ., data=train.df,
                      method="glmnet",  family="binomial",
                      tuneGrid=tuneGrid, trControl=trControl)
coefs <- coef(model$finalModel, s=model$bestTune$lambda)
coefs[which(as.matrix(coefs)!=0),]

confusionMatrix(predict(model, holdout.df), holdout.df$Flight.Status)
```

```{r}
actual <- ifelse(holdout.df$Flight.Status == "delayed", 1, 0)
pred <- predict(model, holdout.df, type="prob")[, 2]
lassoGain <- gains(actual, pred, groups=length(pred))

nactual <-sum(actual)
ggplot() +
  geom_line(aes(x=gain$cume.obs, y=gain$cume.pct.of.total * nactual), color='grey') +
  geom_line(aes(x=redGain$cume.obs, y=redGain$cume.pct.of.total * nactual)) +
  geom_line(aes(x=lassoGain$cume.obs, y=lassoGain$cume.pct.of.total * nactual), color='red') +
  geom_line(aes(x=c(0, max(gain$cume.obs)), y=c(0, nactual)), color="darkgrey") +
  labs(x="# cases", y="Cumulative")

ggsave(file=file.path("..", "figures", "chapter_10", "LRFlightDelaysCoefsReduced.pdf"),
       last_plot() + theme_bw(), width=3, height=3)
```

