---
title: "Machine Learning for Business Analytics"
author: "Chapter 18: Regression-Based Forecasting"
output:
  pdf_document:
    toc: no
    highlight: tango
#  html_document:
#    toc: yes
#    toc_depth: 4
#    toc_float: yes
---
<style>
h1.title { font-size: 28px; }
h1 { font-size: 22px; }
h2 { font-size: 18px; }
h3 { font-size: 14px; }
h4 { font-size: 12px; }
</style>
```{r global_options, include=FALSE}
knitr::opts_chunk$set(error=TRUE,        # Keep compiling upon error
                      collapse=FALSE,    # collapse by default
                      echo=TRUE,         # echo code by default
                      comment = "#>",    # change comment character
                      fig.width = 5.5,     # set figure width
                      fig.align = "center",# set figure position
                      out.width = "49%", # set width of displayed images
                      warning=FALSE,     # do not show R warnings
                      message=FALSE,     # do not show R messages
                      row.print=25)
```


```{r}
if (!require(mlba)) {
  library(devtools)
  install_github("gedeck/mlba/mlba", force=TRUE)
}
options(scipen=999)
```

# A Model with Trend
## Linear Trend
```{r}
library(forecast)
library(ggplot2)
Amtrak.data <- mlba::Amtrak

# create time series
ridership.ts <- ts(Amtrak.data$Ridership,
                   start=c(1991, 1), end=c(2004, 3), freq=12)

# produce linear trend model
ridership.lm <- tslm(ridership.ts ~ trend)

# plot the series using the autoplot function to make use of ggplot
autoplot(ridership.ts, xlab="Time", ylab="Ridership (in 000s)", color="steelblue") +
  autolayer(ridership.lm$fitted.values, color="tomato", size=0.75) +
  scale_y_continuous(limits=c(1300, 2300))
```

```{r}
ggsave(file=file.path("..", "figures", "chapter_18", "AmtrakLinear.pdf"),
         last_plot() + scale_x_continuous(n.breaks=10) + theme_bw(),
         width=6, height=3, units="in")
```

```{r}
nTest <- 36
nTrain <- length(ridership.ts) - nTest

# partition the data
train.ts <- window(ridership.ts, start=c(1991, 1), end=c(1991, nTrain))
test.ts <- window(ridership.ts, start=c(1991, nTrain + 1),
                                   end=c(1991, nTrain + nTest))
```

```{r}
# fit linear trend model to training set
train.lm <- tslm(train.ts ~ trend)
train.lm.pred <- forecast(train.lm, h=length(test.ts), level=0)

# define functions to create consistent plots of models
library(gridExtra)
colData <- "steelblue"; colModel <- "tomato"
plotForecast <- function(model, train.ts, test.ts) {
  model.pred <- forecast(model, h=length(test.ts), level=0)
  g <- autoplot(train.ts, xlab="Time", ylab="Ridership (in 000s)", color=colData) +
    autolayer(test.ts, color=colData) +
    autolayer(model$fitted.values, color=colModel, size=0.75) +
    autolayer(model.pred$mean, color=colModel, size=0.75)
  return (g)
}
plotResiduals <- function(model, test.ts) {
  model.pred <- forecast(model, h=length(test.ts), level=0)
  g <- autoplot(model$residuals, xlab="Time", ylab="Forecast errors", color=colModel, size=0.75) +
    autolayer(test.ts - model.pred$mean, color=colModel, size=0.75) +
    geom_hline(yintercept=0, color="darkgrey") +
    coord_cartesian(ylim=c(-410, 410))
  return (g)
}
g1 <- plotForecast(train.lm, train.ts, test.ts)
g2 <- plotResiduals(train.lm, test.ts)
grid.arrange(g1, g2, nrow=2)
```

```{r}
addTrainValid <- function(g, train.ts, test.ts, yh) {
    delta <- 1/12
    date_t <- time(train.ts)[1]
    date_th <- time(test.ts)[1] - delta
    date_hf <- tail(time(test.ts), 1) + delta
    g <- g + geom_vline(xintercept=date_th, color="darkgrey") +
      geom_segment(aes(x=date_t, xend=date_th-delta, y=yh, yend=yh), color="darkgrey") +
      geom_segment(aes(x=date_th+delta, xend=date_hf-delta, y=yh, yend=yh), color="darkgrey")  +
      geom_text(aes(x=(date_t+date_th)/2, y=yh+50, label='Training')) +
      geom_text(aes(x=(date_th+date_hf)/2, y=yh+50, label='Test'))
    g <- g + scale_x_continuous(n.breaks=10)
    return (g)
  }
  g1 <- addTrainValid(g1, train.ts, test.ts, 2300)
  g2 <- addTrainValid(g2, train.ts, test.ts, 500)# + coord_cartesian(ylim=c(-410, 550))
  gridExtra::grid.arrange(g1, g2, nrow=2)

  g <- arrangeGrob(g1 + theme_bw(), g2 + theme_bw())
  ggsave(file=file.path("..", "figures", "chapter_18", "AmtrakLinearPred.pdf"),
         g, width=6, height=6, units="in")
```

```{r}
summary(train.lm)
accuracy(train.lm.pred, test.ts)
```

## Exponential Trend
```{r}
# fit exponential trend using tslm() with argument lambda = 0
train.lm.expo.trend <- tslm(train.ts ~ trend, lambda=0)
train.lm.expo.trend.pred <- forecast(train.lm.expo.trend, h=nTest, level=0)

# fit linear trend using tslm() with argument lambda = 1 (no transform of y)
train.lm.linear.trend <- tslm(train.ts ~ trend, lambda=1)
train.lm.linear.trend.pred <- forecast(train.lm.linear.trend, h=nTest, level=0)

plotForecast(train.lm.expo.trend, train.ts, test.ts) +
  autolayer(train.lm.linear.trend$fitted.values, color="forestgreen", size=0.75) +
  autolayer(train.lm.linear.trend.pred, color="forestgreen", size=0.75)
```

```{r}
g <- addTrainValid(last_plot(), train.ts, test.ts, 2300) + theme_bw()
  ggsave(file=file.path("..", "figures", "chapter_18", "AmtrakExpoPred.pdf"),
         g, width=6, height=3, units="in")
```

## Polynomial Trend
```{r}
# fit quadratic trend using function I(), which treats an object "as is".
train.lm.poly <- tslm(train.ts ~ trend + I(trend^2))
summary(train.lm.poly)

g1 <- plotForecast(train.lm.poly, train.ts, test.ts)
g2 <- plotResiduals(train.lm.poly, test.ts)
grid.arrange(g1, g2, nrow=2)
```

```{r}
g1 <- addTrainValid(g1, train.ts, test.ts, 2300)
  g2 <- addTrainValid(g2, train.ts, test.ts, 500)# + coord_cartesian(ylim=c(-410, 550))
  grid.arrange(g1, g2, nrow=2)

  g <- arrangeGrob(g1 + theme_bw(), g2 + theme_bw())
  ggsave(file=file.path("..", "figures", "chapter_18", "AmtrakPolyPred.pdf"),
         g, width=6, height=6, units="in")
```

# A Model with Seasonality
```{r}
# fit season model
train.lm.season <- tslm(train.ts ~ season)

g1 <- plotForecast(train.lm.season, train.ts, test.ts)
g2 <- plotResiduals(train.lm.season, test.ts)
grid.arrange(g1, g2, nrow=2)

g1 <- addTrainValid(g1, train.ts, test.ts, 2300)
g2 <- addTrainValid(g2, train.ts, test.ts, 500)
grid.arrange(g1, g2, nrow=2)

g <- arrangeGrob(g1 + theme_bw(), g2 + theme_bw())
ggsave(file=file.path("..", "figures", "chapter_18", "AmtrakSeason.pdf"),
       g, width=6, height=6, units="in")
```

## Additive vs.~Multiplicative Seasonality
```{r}
train.lm.season <- tslm(train.ts ~ season)
summary(train.lm.season)
```

# A Model with Trend and Seasonality
```{r}
train.lm.trend.season <- tslm(train.ts ~ trend + I(trend^2) + season)
g1 <- plotForecast(train.lm.trend.season, train.ts, test.ts)
g2 <- plotResiduals(train.lm.trend.season, test.ts)
grid.arrange(g1, g2, nrow=2)

g1 <- addTrainValid(g1, train.ts, test.ts, 2300)
g2 <- addTrainValid(g2, train.ts, test.ts, 500)
grid.arrange(g1, g2, nrow=2)
g <- arrangeGrob(g1 + theme_bw(), g2 + theme_bw())
ggsave(file=file.path("..", "figures", "chapter_18", "AmtrakTrendSeason.pdf"),
       g, width=6, height=6, units="in")
```

```{r}
summary(train.lm.trend.season)
```

# Autocorrelation and ARIMA Models
## Computing Autocorrelation
```{r}
ridership.24.ts <- window(train.ts, start=c(1991, 1), end=c(1991, 24))
Acf(ridership.24.ts, lag.max=12, main="")
```

```{r}
# ggplot version of graph
  ggAcf(ridership.24.ts, lag.max=12, main="")
  ggsave(file=file.path("..", "figures", "chapter_18", "AmtrakACF.pdf"),
         last_plot() + theme_bw(), width=6, height=3, units="in")
```

```{r}
# ggplot version of graph
  ggAcf(train.lm.trend.season$residuals, lag.max=12, main="")
  ggsave(file=file.path("..", "figures", "chapter_18", "AmtrakResACF.pdf"),
         last_plot() + theme_bw(), width=6, height=3, units="in")
```

## Improving Forecasts by Integrating Autocorrelation Information
```{r}
# fit linear regression with quadratic trend and seasonality to Ridership
train.lm.trend.season <- tslm(train.ts ~ trend + I(trend^2) + season)

# fit AR(1) model to training residuals
# use Arima() in the forecast package to fit an ARIMA model
# (that includes AR models); order = c(1,0,0) gives an AR(1).
train.res.arima <- Arima(train.lm.trend.season$residuals, order = c(1,0,0))
test.res.arima.pred <- forecast(train.res.arima, h = 1)
```

```{r}
summary(train.res.arima)
test.res.arima.pred
```

```{r}
autoplot(train.lm.trend.season$residuals, xlab="Time", ylab="Residuals", color=colData) +
  autolayer(test.res.arima.pred$fitted, color=colModel, size=0.75)
```

```{r}
g <- addTrainValid(last_plot(), train.ts, test.ts, 200) + theme_bw()
  ggsave(file=file.path("..", "figures", "chapter_18", "AmtrakResARIMA.pdf"),
         g, width=6, height=3, units="in")

  g <- ggAcf(train.res.arima$residuals, lag.max=12, main="")
  ggsave(file=file.path("..", "figures", "chapter_18", "AmtrakResARIMAResACF.pdf"),
         g + theme_bw(), width=6, height=3, units="in")
```

## Evaluating Predictability
```{r}
sp500.df <- mlba::SP500
  sp500.ts <- ts(sp500.df$Close, start = c(1995, 5), end = c(2003, 8), freq = 12)
  g <- autoplot(sp500.ts, xlab="Time", ylab="Closing price") + theme_bw()
  ggsave(file=file.path("..", "figures", "chapter_18", "SP500.pdf"),
         g, width=6, height=3, units="in")
```

```{r}
sp500.df <- mlba::SP500
sp500.ts <- ts(sp500.df$Close, start = c(1995, 5), end = c(2003, 8), freq = 12)
delta <- sp500.ts[2:length(sp500.ts)] - sp500.ts[1:length(sp500.ts)-1]
ggAcf(delta, lag.max=12, main="")
```

```{r}
g <- last_plot() + theme_bw()
  ggsave(file=file.path("..", "figures", "chapter_18", "SP500Diff1-ACF.pdf"),
         g, width=6, height=3, units="in")
```

```{r}
sp500.df <- mlba::SP500
sp500.ts <- ts(sp500.df$Close, start = c(1995, 5), end = c(2003, 8), freq = 12)
sp500.arima <- Arima(sp500.ts, order = c(1,0,0))
sp500.arima
```

```{r}
library(zoo)
airTravel.ts <- ts(mlba::Sept11Travel$Air.RPM..000s., start = c(1990, 1), freq = 12)
airTravel.ts <- window(airTravel.ts, start=c(1990, 1), end=c(2001, 8))
g <- autoplot((airTravel.ts - decompose(airTravel.ts)$seasonal) / 1000000,
               xlab="Month",
               ylab="Seasonally adjusted air revenue\n passenger miles ($ billions)") +
  scale_x_yearmon() +
  ylim(0, 63) +
  theme_bw()
g
ggsave(file=file.path("..", "figures", "chapter_18", "seasonalPreSept11air.pdf"),
       g, width=6, height=3)
```

```{r}
toys.ts <- ts(mlba::ToysRUsRevenues$Revenue.in.million..., start=c(1992, 1), freq=4)
autoplot(toys.ts, x="Time", y="Revenue ($ millions)")
g <- last_plot() + theme_bw()
ggsave(file=file.path("..", "figures", "chapter_18", "Exercise-ToysRUs.pdf"),
       last_plot() + theme_bw(), width=5, height=4, units="in")
```

```{r}
n <- length(toys.ts)
train.ts <- window(toys.ts, end=c(1992, n - 2))
test.ts <- window(toys.ts, start=c(1992, n - 1))
summary(tslm(train.ts ~ trend + season))
```

```{r}
walmart <- mlba::WalMartStock
# use zoo to handle the gaps in the dates
walmart.zoo <- zoo(walmart$Close, as.Date(walmart$Date, "%d-%b-%y"))
autoplot(walmart.zoo) +
  labs(x="Time", y="Close price ($)")
g <- last_plot() + theme_bw()
ggsave(file=file.path("..", "figures", "chapter_18", "Exercise-Walmart.pdf"),
       last_plot() + theme_bw(), width=5, height=4, units="in")
```

```{r}
g1 <- ggAcf(walmart$Close, lag.max=12, ylab="Close ACF", main="")
delta <- walmart$Close[2:length(walmart$Close)] - walmart$Close[1:length(walmart$Close)-1]
g2 <- ggAcf(delta, lag.max=12, ylab="Close differences ACF", main="")
grid.arrange(g1, g2, nrow=2)
g <- arrangeGrob(g1 + theme_bw(), g2 + theme_bw())
ggsave(file=file.path("..", "figures", "chapter_18", "Exercise-Walmart-acf.pdf"),
       g, width=5, height=5, units="in")
```

```{r}
sales.ts <- ts(mlba::DepartmentStoreSales$Sales, freq=4)
autoplot(sales.ts, xlab="Year-quarter", ylab="Sales") +
  geom_point() +
  scale_x_yearqtr(format = "Y%Y-Q%q")
ggsave(file=file.path("..", "figures", "chapter_18", "Exercise-DeptStore.pdf"),
       last_plot() + theme_bw(), width=6, height=3, units="in")
```

```{r}
train.ts <- window(sales.ts, end=c(1, 20))
test.ts <- window(sales.ts, start=c(1, 21))
model <- tslm(train.ts ~ trend + season, lambda=0)
summary(model)
```

```{r}
g1 <- autoplot(train.ts, xlab="Year", ylab="Sales") +
  geom_point() +
  autolayer(model$fitted.values, color=colModel) +
  scale_y_continuous(labels = scales::comma)
g2 <- autoplot(model$residuals, xlab="Year", ylab="Residuals")
grid.arrange(g1, g2, nrow=2)
g <- arrangeGrob(g1 + theme_bw(), g2 + theme_bw())
ggsave(file=file.path("..", "figures", "chapter_18", "Exercise-DeptStore-model.pdf"),
       g, width=5, height=5, units="in")
```

```{r}
sales.df <- mlba::SouvenirSales
sales.ts <- ts(sales.df$Sales, start=c(1995, 1), end=c(2001, 12), freq=12)

g1 <- autoplot(sales.ts, xlab="Time", ylab="Sales (Australian $)") +
  geom_point(size=0.5) +
  scale_y_continuous(labels = scales::comma)
g2 <- g1 + 
  scale_y_log10(labels = scales::comma)
grid.arrange(g1, g2, ncol=2)
g <- arrangeGrob(g1 + theme_bw(), g2 + theme_bw(), ncol=2)
ggsave(file=file.path("..", "figures", "chapter_18", "Exercise-SouvenirShop.pdf"),
       g, width=8, height=4, units="in")
```

```{r}
shipments.df <- mlba::ApplianceShipments
shipments.ts <- ts(shipments.df$Shipments, start=c(1985, 1), freq=4)

autoplot(shipments.ts, xlab="Time", ylab="Shipments (in $000s)") +
  geom_point(size=0.5)
g <- last_plot() + theme_bw()
ggsave(file=file.path("..", "figures", "chapter_18", "Exercise-Appliances.pdf"),
       g, width=5, height=2.5, units="in")
```

```{r}
wines <- mlba::AustralianWines
wines.ts <- ts(wines, start=c(1980, 1), freq=12)

g1 <- autoplot(wines.ts[, "sweet.white"], main="Sweet wine sales",
               xlab="Year", ylab="Thousands of liters")
g2 <- autoplot(wines.ts[, "rose"], main="Rose wine sales",
               xlab="Year", ylab="Thousands of liters")
g3 <- autoplot(wines.ts[, "sparkling"], main="Sparkling wine sales",
               xlab="Year", ylab="Thousands of liters")
g4 <- autoplot(wines.ts[, "red"], main="Red wine sales",
               xlab="Year", ylab="Thousands of liters")
g5 <- autoplot(wines.ts[, "dry.white"], main="Dry white wine sales",
               xlab="Year", ylab="Thousands of liters")
g6 <- autoplot(wines.ts[, "fortified"], main="Fortified wine sales",
               xlab="Year", ylab="Thousands of liters")

grid.arrange(g1, g2, g3, g4, g5, g6, ncol=2, nrow=3)
g <- arrangeGrob(g1 + theme_bw(), g2 + theme_bw(),
                 g3 + theme_bw(), g4 + theme_bw(),
                 g5 + theme_bw(), g6 + theme_bw(),
                 ncol=2, nrow=3)
ggsave(file=file.path("..", "figures", "chapter_18", "Exercise-Wines.pdf"),
       g, width=7, height=9, units="in")
```

